---
title: "Analysis of Factors Affecting Fuel Efficiency"
author: "Graham King"
date: "February 2015"
output:
  html_document:
    css: report.css
---
```{r opts, echo=FALSE}
library(knitr)
opts_chunk$set(comment = NA, fig.width = 12, fig.height = 6, echo = FALSE, size = 'small')

read_chunk("init.R")
read_chunk("mpg.R")
```
###Overview
In this report, I will look at the effect that the weight of a car has on its fuel economy, and what impact the transmission has on this.

```{r init}
<<init>>
```
### Content
```{r subset}
<<subset>>
```

AN initial look at the complete dataset shows that there is a clear linear relationship between car weight and fuel economy.
```{r plot1}
<<plot1>>
```

The sum of all residuals gives an indication of variance:
```{r model1}
<<model1>>
```

When looking at Automatic and Manual transmission data separately, there is a distinct difference in the regression line:
```{r plot2}
<<plot2>>
```

This figure (N) indicates that fuel economy (mpg) reduces faster for manual cars as the weight increases, although at lower weights, the fuel economy is superior than for automatic cars.

We can demonstrate that this model is a better fit by looking at the sum of all residuals for this data:

```{r model2}
<<model2>>
```

This variance(??) is lower than for the combined dataset.

The regression model coefficients are as follows:

```{r model2}
<<model2>>
```

Specifically, for each 1,000lb (??) increase in weight, the manual cars reduce mpg by xyz, compared with abc for automatic cars.

We can use this model to predict weights for hypothetical cars:

```{r predictions}
<<predictions>>
```

Under this model, a hypothetical manual transmission car weighing 1lbs would have a higher mpg than an automatic car of the same weight. Conversely, a hypothetical automatic car weighing 4lbs would have a l

****

### Appendix
Code1
```{r ap1, eval = FALSE, echo = TRUE}
<<subset>>
```